---
title: "Turtle Strandings"
author: "Florian Mayer (DBCA) and Erina Young (Murdoch)"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(tidyverse)
library(lubridate)
library(magrittr)
library(skimr)
library(leaflet)
library(mapview)
# library(ckanr)
```

# Mapping data in QGIS
Open a browser, login to a DBCA site so that your browser session is authenticated.
Then save [this URL](https://strandings.dpaw.wa.gov.au/api/1/animal-encounters/?encounter_type=stranding&taxon=Cheloniidae&format=json)
as e.g. `strandings.geojson` and open it in Quantum GIS.

# Load data
Data are retrieved from [WAStD](https://strandings.dpaw.wa.gov.au/) and parsed
using the R package [wastdr](https://parksandwildlife.github.io/wastdr/).

Consult the excellent [RStudio cheatsheets](https://www.rstudio.com/resources/cheatsheets/) 
son how to use RStudio Desktop and the R packages we use here.

```{r load_data, echo=T}
if (file.exists("~/ts_records.Rda")){
    load("~/ts_records.Rda")
} else {
    query <- list(taxon='Cheloniidae', encounter_type='stranding', format='json')
    ts_records <- wastdr::get_wastd("animal-encounters", query=query)
    save(ts_records, file = "~/ts_records.Rda")
    load("~/ts_records.Rda")
    # listviewer::jsonedit(utils::head(ts_records$features))
}
strandings <- wastdr::parse_animal_encounters(ts_records)
save(strandings, file = "~/strandings.Rda")
```

# Helpers
A few helper functions are defined, but hidden. 
Set `echo=T` in the following chunk to show.

```{r filter_data, echo=F}
species_colours <- tibble::tibble(
    species = c(
    "cheloniidae-fam",
    "chelonia-mydas",
    "eretmochelys-imbricata",
    "natator-depressus",
    "corolla-corolla",
    "lepidochelys-olivacea",
    "caretta-caretta"    
    ),
    species_colours = c(
    "gray",
    "green",
    "darkblue",
    "beige",
    "pink",
    "darkgreen",
    "orange"
    )
)


add_lookups <- . %>% left_join(species_colours, by="species")

filter_2016ff <- . %>% dplyr::filter(date > dmy("31/12/2015"))
filter_2016 <- . %>% dplyr::filter(date > dmy("31/12/2015"), date < dmy("01/01/2017"))
filter_2017 <- . %>% dplyr::filter(date > dmy("31/12/2016"), date < dmy("01/01/2018"))
filter_2018 <- . %>% dplyr::filter(date > dmy("31/12/2017"), date < dmy("01/01/2019"))

filter_broome <- . %>% dplyr::filter(area_name=="Cable Beach Broome")
filter_eighty_mile_beach <- . %>% dplyr::filter(area_name=="Eighty Mile Beach Caravan Park")
filter_anna_plains <- . %>% dplyr::filter(area_name=="Anna Plains")
filter_port_hedland <- . %>% dplyr::filter(site_name=="Port Hedland Turtle Nesting Beaches")
filter_west_pilbara <- . %>% dplyr::filter(area_name=="West Pilbara Turtle Program beaches Wickam")
filter_delambre <- . %>% dplyr::filter(area_name=="Delambre Island")
filter_rosemary <- . %>% dplyr::filter(area_name=="Rosemary Island")
filter_thevenard <- . %>% dplyr::filter(area_name=="Thevenard Island")

strandings_map <- function(data) {
    l <- leaflet(width=800, height=600) %>% 
        addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
        addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
        clearBounds()

    data.df <-  data %>% split(data$species)
    
    names(data.df) %>%
        purrr::walk( function(df) {
            l <<- l %>%
                addAwesomeMarkers(
                    data = data.df[[df]],
                    lng = ~longitude, lat=~latitude,
                    icon = leaflet::makeAwesomeIcon(
                        # text = ~species,
                        markerColor = ~species_colours),
                    label=~paste(date, health, species),
                    popup=~paste(date, sex, health, species, name),
                    group = df
                )
        })
    
    l %>%
        addLayersControl(
            baseGroups = c("Aerial", "Place names"),
            overlayGroups = names(data.df),
            options = layersControlOptions(collapsed = FALSE)
        )
}
```

# Maps

## Strandings 2016

```{r map_strandings_2016, fig.height=5, fig.width=7}
strandings %>% add_lookups %>% filter_2016 %>% strandings_map %T>% mapshot(., file = "strandings_2016.png")
```

## Strandings 2017

```{r map_strandings_2017, fig.height=5, fig.width=7}
strandings %>% add_lookups %>% filter_2017 %>% strandings_map %T>% mapshot(., file = "strandings_2017.png")
```

# Inspect data

```{r}
strandings %>% filter_2016ff %>% skim
strandings %>% filter_2016ff %>% select(-obs) %>% DT::datatable()
```

# Export
```{r save_csv}
strandings %>% filter_2016ff %>% select(-obs) %>% write_delim("strandings.csv", delim = ",")
```
